Perturbing SU($n$)-symmetric interactions

### 2020-01-16 collective_sun/sun_perturbation.tex
some improvements to text, removing the implicit assumption of dimension D=1 by default

To compute the effective Hamiltonian induced by the single-body
perturbation $\V_1$, we use the permutational symmetry of the
ground-state manifold to simplify
\begin{align}
  H_1^{(1)} = \sum_{X,p} v_{Xp} \P_0 X_p \P_0
  \EQFS \sum_{X,p} v_{Xp} X_0
  \EQFS \sum_X \bar v_X \col{X},
  \label{eq:H_1_1}
\end{align}
where $\EQFS$ denotes equality under a restriction to the fully
symmetric manifold $\M_0$, and
\begin{align}
  \bar v_X \equiv \f1N \sum_p v_{Xp},
  &&
  \col{X} \equiv \sum_p X_p,
\end{align}
are respectively the mean value in $\v v_X$ and the collective version
of $X$.  In order to compute the second order effective Hamiltonian
$H_1^{(2)}$, we first choose an arbitrary fully symmetric state
$\ket\psi\in\M_0$ and expand (see Appendix \ref{sec:H_V1_psi}):
\begin{align}
  H_0 \V_1 \ket\psi
  = E_0 \V_1 \ket\psi
  + \sum_X \p{\sum_{p,q} h_{pq} v_{Xp} X_q - \sum_p h_p v_{Xp} X_p}
  \ket\psi,
  &&
  h_p \equiv \sum_q h_{pq},
  \label{eq:H_V1_psi}
\end{align}
where we define $h_{pq}$ for all $p,q$ (as opposed to only $p<q$) by
enforcing $h_{pq}=h_{qp}$ and $h_{pp}=0$.  We thus find that a
perturbation $\V_1^\Delta\ket\psi$ indexed by a scalar $\Delta$ is an
eigenvector of the interaction Hamiltonian $H_0$ if its coefficient
vectors $\v v_X^\Delta$ all satisfy the eigenvalue equation
\begin{align}
  \p{\m h - \diag\v h}\c\v v = \Delta \v v,
  &&
  \m h \equiv \sum_{p,q} h_{pq} \op{p}{q},
  &&
  \v h \equiv \sum_q h_q \ket{q},
  \label{eq:cond_1}
\end{align}
where $\diag\v h\equiv\sum_p h_p \op{p}$ is a matrix with $\v h$ on
the diagonal and zeros everywhere else.  Solving the general
eigenvalue problem in \eqref{eq:cond_1} requires diagonalizing the
$N\times N$ matrix $\m h-\diag\v h$.  When the interaction Hamiltonian
$H_0$ is translationally invariant, however, this eigenvalue problem
can be solved analytically; we provide the corresponding solution in
Appendix \ref{sec:trans_inv}.

@@

The first-order effective Hamiltonian $H_2^{(1)}$ induced on the fully
symmetric manifold $\M_0$ by the two-body perturbation $\V_2$ in
\eqref{eq:perturbations} is simply the restriction of $\V_2$ onto
$\M_0$:
\begin{align}
  H_2^{(1)} = \sum_O \sum_{p<q} w_{Opq} \P_0 O_{pq} \P_0
  \EQFS \sum_O \sum_{p<q} w_{Opq} O_{0,1}
  \EQFS \sum_O \bar w_O \col{O},
\end{align}
where
\begin{align}
  \bar w_O \equiv {N \choose 2}^{-1} \sum_{p<q} w_{Opq},
  &&
  \col{O} \equiv \sum_{p<q} O_{pq},
\end{align}
are an average of $w_{Opq}$ and a collective version of the two-body
operator $O_{pq}$.  If the two-body operator $O=X\otimes Y$, then
\begin{align}
  H_2^{(1)} \EQFS \f12 \sum_{\p{X,Y}} \bar w_{\p{X,Y}}
  \p{\col{X}\,\col{Y} - \col{XY}}.
\end{align}
In order to compute the second-order effective Hamiltonian
$H_2^{(2)}$, as before we pick an arbitrary fully symmetric state
$\ket\psi\in\M_0$ and expand (see Appendix \ref{sec:H_V2_psi})
\begin{align}
  H_0 \V_2 \ket\psi
  = E_0 \V_2 \ket\psi
  + \f12 \sum_{p\ne q} \sp{2\bar h_{pq} w_{Opq}
    + \sum_k \p{h_{pk} w_{Okq} + w_{Opk} h_{kq}}}
  O_{pq} \ket\psi,
  \label{eq:H_V2_psi}
\end{align}
where to keep notation compact, we define
\begin{align}
  \bar h_{pq} \equiv h_{pq} - \f12\p{h_p + h_q},
  &&
  h_p \equiv \sum_q h_{pq},
\end{align}
and we define $w_{Opq}$ for all $p,q$ (as opposed to only $p<q$) by
enforcing $w_{Opq}=w_{Oqp}$ and $w_{Opp}=0$.  We thus find that the
vector $\V_2^\Delta\ket\psi$ indexed by a scalar $\Delta$ is an
eigenvector of the interaction Hamiltonian $H_0$ if its coefficients
$\v{\m w}_O^\Delta$ all satisfy the eigenvalue equation
\begin{align}
  \check{\m h} \c \v{\m w} = \Delta \v{\m w},
  \label{eq:cond_2}
\end{align}
where
\begin{align}
  \check{\m h} \equiv 2\sum_{p<q} \bar h_{pq} \op{\set{p,q}}
  + \sum_{p<q} \ket{\set{p,q}} \sum_{k\notin\set{p,q}}
  \p{h_{pk} \bra{\set{k,q}} + h_{kq} \bra{\set{p,k}}}.
  \label{eq:h_super_mat}
\end{align}
The eigenvalue problem in \eqref{eq:cond_2} can be simplified further
when the interaction Hamiltonian $H_0$ is translationally invariant;
we discuss this simplification in Appendix \ref{sec:trans_inv}.

@@

Here we simplify the eigenvalue problems that appear at second order
in the single- and two-body theory in Sections
\ref{sec:single_body_pert} and \ref{sec:two_body_pert} in the case of
a translationally invariant system with periodic boundary conditions.
We solve the single-body eigenvalue problem analytically, and reduce
the two-body eigenvalue problem to that of diagonalizing an
$\p{N-1}\times\p{N-1}$ matrix.  Finally, we discuss the reduction of
the two-body problem to that of diagonalizing a
$\sim\p{N/2^D}\times\p{N/2^D}$ matrix, and work out the case of $D=1$
explicitly as an example.

@@

For reference, the single-body eigenvalue problem is
\begin{align}
  \p{\m h - \diag\v h}\c\v v = \Delta \v v,
  \label{eq:cond_1_ref}
\end{align}
where $\m h$ is a matrix of all couplings $h_{pq}$ that appear in the
interaction Hamiltonian $H_0$ in \eqref{eq:ints}; $\v h$ is a vector
of all $h_p\equiv\sum_q h_{pq}$; $\diag\v h$ is a matrix with $\v h$
on the diagonal and zeroes everywhere else; and $\v v$ is a vector of
coefficients used to construct the single-body perturbation
$\V_1\p{\v v}$ defined in \eqref{eq:perturbations}.  When the
interaction Hamiltonian $H_0$ is translationally invariant, the
couplings $h_{pq}$ depend only on the separation $\abs{p-q}$, so the
eigenvectors of $\m h$ are plane waves of the form
\begin{align}
  \v v_k \equiv \sum_p e^{ipk} \ket{p},
\end{align}
where $k\in\ZZ_N\times2\pi/N$ (with
$\ZZ_N\equiv\set{0,1,\cdots,N-1}$).  The corresponding eigenvalues can
be determined by expanding
\begin{align}
  \m h\c\v v_k = \sum_{p,q} h_{pq} e^{iqk} \ket{p}
  = \sum_{p,d} h_{p,d+p} e^{i\p{d+p}k} \ket{p}
  = \sum_d h_{0,d} e^{idk} \v v_k
  = \sum_d h_{0,d} \cos\p{dk} \v v_k,
\end{align}
where the imaginary terms vanish due to the fact that
$h_{0,d}=h_{0,-d}$.  The remainder of \eqref{eq:cond_1_ref} that we
need to sort out is $\diag\v h$, where all
$h_p=\sum_qh_{pq}=\sum_dh_{0,d}$ are equal, which implies that
$\diag\v h=\sum_dh_{0,d}$ is a scalar.  We thus find that
\begin{align}
  \p{\m h - \diag\v h}\c\v v_k = \Delta_k \v v_k,
  &&
  \Delta_k \equiv \sum_d h_{0,d} \sp{\cos\p{dk}-1}.
\end{align}
This solution also holds for a $D$-dimensional periodic lattice of
$N=L^D$ spins, with $d\in\ZZ_L^D$ and $k\in\ZZ_L^D\times 2\pi/L$ (and
$dk\to d\c k\equiv\sum_{j=1}^D d_j k_j$).

@@

The two-body eigenvalue equation in \eqref{eq:cond_2} is designed to
find solutions $\m w$ to the equation
\begin{align}
  \p{2h_{pq}-h_p-h_q} w_{pq}
  + \sum_{k\notin\set{p,q}} \p{h_{pk} w_{kq} + w_{pk} h_{kq}}
  = \Delta w_{pq},
  \label{eq:cond_2_ref_start}
\end{align}
where $h_p\equiv\sum_q h_{pq}$.  If the interaction Hamiltonian $H_0$
and perturbation $\V_2\p{\m w}$ are translationally invariant, then
the couplings $h_{pq}$ and coefficients $w_{pq}$ depend on $p$ and $q$
only through the separation $\abs{p-q}$.  In this case, all $h_p=h_0$
are equal, and \eqref{eq:cond_2_ref_start} can be written in the form
\begin{align}
  2 \m h\circ \m w + \m h\c\m w + \m w\c\m h
  = \sp{\Delta + 2h_0} \m w,
  \label{eq:cond_2_ref}
\end{align}
where $\m h\circ\m w$ denotes an element-wise (Kronecker) product of
$\m h$ and $\m w$, i.e.~with matrix elements $h_{pq}w_{pq}$.  In order
to find solutions to \eqref{eq:cond_2_ref}, we we expand the matrices
$\m h,\m w$ in a basis that respects translational symmetry.  To this
end, for all $d\in\ZZ_N$ we define the {\it shift matrices}
\begin{align}
  \m s_d \equiv \sum_p \op{p+d}{p},
\end{align}
where the sum $p+d$ is implicitly taken modulo $N$.  The shift
matrices allow us to expand
\begin{align}
  \m h = \sum_{p,q} h_{pq} \op{p}{q}
  = \sum_{d,q} h_{q+d,q} \op{q+d}{q}
  = \sum_d h_{d,0} \m s_d,
  &&
  \m w = \sum_d w_{d,0} \m s_d,
\end{align}
and write the condition in \eqref{eq:cond_2_ref} as
\begin{align}
  \sum_{d\ne 0} h_{d,0} w_{d,0} \m s_d
  + \sum_{\substack{c,d\ne0\\c+d\ne0}} h_{c,0} w_{d,0} \m s_{c+d}
  = \p{\f12\Delta + h_0} \sum_{d\ne 0} w_{d,0} \m s_d,
\end{align}
where we explicitly exclude terms that always vanish due to the fact
that $\m h,\m w$ are zero on the diagonal.  With some re-indexing, we
can write
\begin{align}
  \sum_{d\ne 0} h_{d,0} w_{d,0} \m s_d
  + \sum_{\substack{c,d\ne0}} h_{d-c,0} w_{c,0} \m s_d
  = \p{\f12\Delta + h_0} \sum_{d\ne 0} w_{d,0} \m s_d,
  \label{eq:cond_2_shift}
\end{align}
and define
\begin{align}
  \ut{\check{\m h}}
  \equiv \sum_{d\ne0} h_{d,0} \op{d}
  + \sum_{c,d\ne0} h_{d-c,0} \op{d}{c},
  &&
  \ut{\v{\m w}} \equiv \sum_{d\ne0} w_{d,0} \ket{d},
\end{align}
in order to express \eqref{eq:cond_2_shift} as the eigenvalue
equation
\begin{align}
  \ut{\check{\m h}} \c \ut{\v{\m w}}
  = \p{\f12\Delta + h_0} \ut{\v{\m w}}.
  \label{eq:cond_2_square}
\end{align}
This eigenvalue problem takes an identical form a $D$ dimensional
periodic lattice of $N=L^D$ spins, in which case lattice sites and
shift matrices are indexed by vectors $d\in\ZZ_L^D$.

We can reduce the eigenvalue problem in \eqref{eq:cond_2_square} even
further by making use of reflection symmetries such as
$h_{d,0}=h_{-d,0}$.  In $D$ dimensions, there is one such symmetry for
each coordinate that indexes lattice sites, which implies that the
degrees of freedom in $\ut{\v{\m w}}$ gets reduced by a factor of
$\sim 2^D$.  Reflection symmetries similarly reduce the degrees of
freedom in $\ut{\check{\m h}}$ from $\sim N\times N$ to
$\sim\p{N/2^D}\times\p{N/2^D}$.  We work out the case of $D=1$ below
as an example, and leave the general case of arbitrary dimension
$D\ge1$ to the future, if and when it becomes necessary.

>>>>

To compute the effective Hamiltonian induced by the single-body
perturbation $\V_1$, we use the permutational symmetry of the
ground-state manifold to simplify
\begin{align}
  H_1^{(1)} = \sum_{X,p} v_{Xp} \P_0 X_p \P_0
  \EQFS \sum_{X,p} v_{Xp} X_0
  \EQFS \sum_X \bar v_X \col{X},
  \label{eq:H_1_1}
\end{align}
where $\EQFS$ denotes equality under a restriction to the fully
symmetric manifold $\M_0$, and
\begin{align}
  \bar v_X \equiv \f1N \sum_p v_{Xp},
  &&
  \col{X} \equiv \sum_p X_p,
\end{align}
are respectively the mean value in $\v v_X$ and the collective version
of $X$.  In order to compute the second order effective Hamiltonian
$H_1^{(2)}$, we first choose an arbitrary fully symmetric state
$\ket\psi\in\M_0$ and expand (see Appendix \ref{sec:H_V1_psi}):
\begin{align}
  H_0 \V_1 \ket\psi
  = E_0 \V_1 \ket\psi
  + \sum_X \p{\sum_{p,q} h_{pq} v_{Xp} X_q - \sum_p h_p v_{Xp} X_p}
  \ket\psi,
  &&
  h_p \equiv \sum_q h_{pq},
  \label{eq:H_V1_psi}
\end{align}
where we define $h_{pq}$ for all $p,q$ (as opposed to only $p<q$) by
enforcing $h_{pq}=h_{qp}$ and $h_{pp}=0$.  We thus find that a
perturbation $\V_1^\Delta\ket\psi$ indexed by a scalar $\Delta$ is an
eigenvector of the interaction Hamiltonian $H_0$ if its coefficient
vectors $\v v_X^\Delta$ all satisfy the eigenvalue equation
\begin{align}
  \p{\m h - \diag\v h}\c\v v = \Delta \v v,
  &&
  \m h \equiv \sum_{p,q} h_{pq} \op{p}{q},
  &&
  \v h \equiv \sum_q h_q \ket{q},
  \label{eq:cond_1}
\end{align}
where $\m h$ is a matrix of all $h_{pq}$, and
$\diag\v h\equiv\sum_p h_p \op{p}$ is a matrix with $\v h$ on the
diagonal and zeros everywhere else.  Solving the general eigenvalue
problem in \eqref{eq:cond_1} requires diagonalizing the $N\times N$
matrix $\m h-\diag\v h$.  When the interaction Hamiltonian $H_0$ is
translationally invariant, however, this eigenvalue problem can be
solved analytically; we provide the corresponding solution in Appendix
\ref{sec:trans_inv}.

@@

The first-order effective Hamiltonian $H_2^{(1)}$ induced on the fully
symmetric manifold $\M_0$ by the two-body perturbation $\V_2$ in
\eqref{eq:perturbations} is simply the restriction of $\V_2$ onto
$\M_0$:
\begin{align}
  H_2^{(1)} = \sum_O \sum_{p<q} w_{Opq} \P_0 O_{pq} \P_0
  \EQFS \sum_O \sum_{p<q} w_{Opq} O_{0,1}
  \EQFS \sum_O \bar w_O \col{O},
\end{align}
where
\begin{align}
  \bar w_O \equiv {N \choose 2}^{-1} \sum_{p<q} w_{Opq},
  &&
  \col{O} \equiv \sum_{p<q} O_{pq},
\end{align}
are an average of $w_{Opq}$ and a collective version of the two-body
operator $O_{pq}$.  If the two-body operator $O=X\otimes Y$, then
\begin{align}
  H_2^{(1)} \EQFS \f12 \sum_{\p{X,Y}} \bar w_{\p{X,Y}}
  \p{\col{X}\,\col{Y} - \col{XY}}.
\end{align}
In order to compute the second-order effective Hamiltonian
$H_2^{(2)}$, as before we pick an arbitrary fully symmetric state
$\ket\psi\in\M_0$ and expand (see Appendix \ref{sec:H_V2_psi})
\begin{align}
  H_0 \V_2 \ket\psi
  = E_0 \V_2 \ket\psi
  + \f12 \sum_{p\ne q} \sp{2\bar h_{pq} w_{Opq}
    + \sum_k \p{h_{pk} w_{Okq} + w_{Opk} h_{kq}}}
  O_{pq} \ket\psi,
  \label{eq:H_V2_psi}
\end{align}
where to keep notation compact, we define
\begin{align}
  \bar h_{pq} \equiv h_{pq} - \f12\p{h_p + h_q},
  &&
  h_p \equiv \sum_q h_{pq},
\end{align}
and we define $w_{Opq}$ for all $p,q$ (as opposed to only $p<q$) by
enforcing $w_{Opq}=w_{Oqp}$ and $w_{Opp}=0$.  We thus find that the
vector $\V_2^\Delta\ket\psi$ indexed by a scalar $\Delta$ is an
eigenvector of the interaction Hamiltonian $H_0$ if its coefficients
$\v{\m w}_O^\Delta$ all satisfy the eigenvalue equation
\begin{align}
  \check{\m h} \c \v{\m w} = \Delta \v{\m w},
  \label{eq:cond_2}
\end{align}
where
\begin{align}
  \check{\m h} \equiv 2\sum_{p<q} \bar h_{pq} \op{\set{p,q}}
  + \sum_{p<q} \ket{\set{p,q}} \sum_{k\notin\set{p,q}}
  \p{h_{pk} \bra{\set{k,q}} + h_{kq} \bra{\set{p,k}}}.
  \label{eq:h_super_mat}
\end{align}
Solving the general eigenvalue problem in \eqref{eq:cond_2} requires
diagonalizing the ${N \choose 2}\times{N \choose 2}\sim N^2\times N^2$
matrix $\check{\m h}$.  When the interaction Hamiltonian $H_0$ is
translationally invariant, however, this eigenvalue problem can be
reduced to that of diagonalizing an $\sim N\times N$ matrix; we
discuss this reduction in Appendix \ref{sec:trans_inv}.

@@

Here we simplify the eigenvalue problems that appear at second order
in the single- and two-body theory in Sections
\ref{sec:single_body_pert} and \ref{sec:two_body_pert} in the case of
a translationally invariant system with periodic boundary conditions.
We solve the single-body eigenvalue problem analytically, and reduce
the two-body eigenvalue problem to that of diagonalizing an
$\p{N-1}\times\p{N-1}$ matrix.  Finally, we discuss the further
reduction of the two-body problem to that of diagonalizing a matrix
with dimensions $\sim\p{N/2^D}\times\p{N/2^D}$ in the case of
$D$-dimensional periodic lattice, and work out the case of $D=1$
explicitly as an example.

@@

For reference, the single-body eigenvalue problem is
\begin{align}
  \p{\m h - \diag\v h}\c\v v = \Delta \v v,
  \label{eq:cond_1_ref}
\end{align}
where $\m h$ is a matrix of all couplings $h_{pq}$ that appear in the
interaction Hamiltonian $H_0$ in \eqref{eq:ints}; $\v h$ is a vector
of all $h_p\equiv\sum_q h_{pq}$; $\diag\v h$ is a matrix with $\v h$
on the diagonal and zeroes everywhere else; and $\v v$ is a vector of
coefficients used to construct the single-body perturbation
$\V_1\p{\v v}$ defined in \eqref{eq:perturbations}.  When the
interaction Hamiltonian $H_0$ is translationally invariant, the
couplings $h_{pq}$ depend only on the separation $\abs{p-q}$, so the
eigenvectors of $\m h$ are plane waves of the form
\begin{align}
  \v v_k \equiv \sum_p e^{ip\c k} \ket{p},
\end{align}
where on a $D$-dimensional periodic lattice of $N=L^D$ spins, lattice
sites are indexed by vectors $p\in\ZZ_L^D$, and wavenumbers take on
values $k\in\ZZ_L^D\times2\pi/N$.  The corresponding eigenvalues of
$\m h$ can be determined by expanding
\begin{align}
  \m h\c\v v_k = \sum_{p,q} h_{pq} e^{iq\c k} \ket{p}
  = \sum_{p,d} h_{p,d+p} e^{i\p{d+p}\c k} \ket{p}
  = \sum_d h_{0,d} e^{id\c k} \v v_k
  = \sum_d h_{0,d} \cos\p{d\c k} \v v_k,
\end{align}
where the imaginary terms vanish due to the fact that
$h_{0,d}=h_{0,-d}$.  The remainder of \eqref{eq:cond_1_ref} that we
need to sort out is $\diag\v h$, where all
$h_p=\sum_qh_{pq}=\sum_dh_{0,d}$ are equal, which implies that
$\diag\v h=\sum_dh_{0,d}$ is a scalar.  We thus find that
\begin{align}
  \p{\m h - \diag\v h}\c\v v_k = \Delta_k \v v_k,
  &&
  \Delta_k \equiv \sum_d h_{0,d} \sp{\cos\p{d\c k}-1}.
\end{align}

@@

The two-body eigenvalue equation in \eqref{eq:cond_2} is designed to
find solutions $\m w$ to the equation
\begin{align}
  \p{2h_{pq}-h_p-h_q} w_{pq}
  + \sum_{k\notin\set{p,q}} \p{h_{pk} w_{kq} + w_{pk} h_{kq}}
  = \Delta w_{pq},
  \label{eq:cond_2_ref_start}
\end{align}
where $h_p\equiv\sum_q h_{pq}$.  If the interaction Hamiltonian $H_0$
and perturbation $\V_2\p{\m w}$ are translationally invariant, then
the couplings $h_{pq}$ and coefficients $w_{pq}$ depend on $p$ and $q$
only through the separation $\abs{p-q}$.  In this case, all $h_p=h_0$
are equal, and \eqref{eq:cond_2_ref_start} can be written in the form
\begin{align}
  2 \m h\circ \m w + \m h\c\m w + \m w\c\m h
  = \sp{\Delta + 2h_0} \m w,
  \label{eq:cond_2_ref}
\end{align}
where $\m h$ and $\m w$ are matrices of all $h_{pq}$ and $w_{pq}$, and
$\m h\circ\m w$ denotes an element-wise (Kronecker) product of $\m h$
and $\m w$, i.e.~with matrix elements $h_{pq}w_{pq}$.  In order to
find solutions to \eqref{eq:cond_2_ref}, we we expand the matrices
$\m h,\m w$ in a basis that respects translational symmetry.  To this
end, for all lattice site displacements $d\in\ZZ_L^D$ we define the
{\it shift matrices}
\begin{align}
  \m s_d \equiv \sum_p \op{p+d}{p},
\end{align}
where the element-wise sums in $p+d$ are implicitly taken modulo $L$
(i.e.~the length of the periodic lattice along each axis).  The shift
matrices allow us to expand
\begin{align}
  \m h = \sum_{p,q} h_{pq} \op{p}{q}
  = \sum_{d,q} h_{q+d,q} \op{q+d}{q}
  = \sum_d h_{d,0} \m s_d,
  &&
  \m w = \sum_d w_{d,0} \m s_d,
\end{align}
and write the condition in \eqref{eq:cond_2_ref} as
\begin{align}
  \sum_{d\ne 0} h_{d,0} w_{d,0} \m s_d
  + \sum_{\substack{c,d\ne0\\c+d\ne0}} h_{c,0} w_{d,0} \m s_{c+d}
  = \p{\f12\Delta + h_0} \sum_{d\ne 0} w_{d,0} \m s_d,
\end{align}
where we explicitly exclude terms that always vanish due to the fact
that $\m h,\m w$ are zero on the diagonal.  With some re-indexing, we
can write
\begin{align}
  \sum_{d\ne 0} h_{d,0} w_{d,0} \m s_d
  + \sum_{\substack{c,d\ne0}} h_{d-c,0} w_{c,0} \m s_d
  = \p{\f12\Delta + h_0} \sum_{d\ne 0} w_{d,0} \m s_d,
  \label{eq:cond_2_shift}
\end{align}
and define
\begin{align}
  \ut{\check{\m h}}
  \equiv \sum_{d\ne0} h_{d,0} \op{d}
  + \sum_{c,d\ne0} h_{d-c,0} \op{d}{c},
  &&
  \ut{\v{\m w}} \equiv \sum_{d\ne0} w_{d,0} \ket{d},
\end{align}
in order to express \eqref{eq:cond_2_shift} as the eigenvalue
equation
\begin{align}
  \ut{\check{\m h}} \c \ut{\v{\m w}}
  = \p{\f12\Delta + h_0} \ut{\v{\m w}}.
  \label{eq:cond_2_square}
\end{align}
We can reduce this eigenvalue problem even further by making use of
reflection symmetries such as $h_{d,0}=h_{-d,0}$.  In $D$ dimensions,
there is one such symmetry for each lattice axis, which implies that
the degrees of freedom in $\ut{\v{\m w}}$ gets reduced by a factor of
$\sim 2^D$.  Reflection symmetries similarly reduce the degrees of
freedom in $\ut{\check{\m h}}$ from $\sim N\times N$ to
$\sim\p{N/2^D}\times\p{N/2^D}$.  We work out the case of $D=1$ below
as an example, and leave the general case of arbitrary dimension
$D\ge1$ to future work.
